---
title: "ClipPC_2_shp"
author: "Abhinav Shrestha"
date: "2022-11-05"
output: html_document
---

###################################################
# notes

Script clips point cloud data (.las) to shapefiles  

###################################################

##Loading required packages
```{r}
require(sp)
require(sf)
require(raster)
require(rLiDAR)
require(lidR)
require(rgdal)
require(terra)

```

##Importing point cloud as LAS catalog 
```{r}

las <- readLAScatalog("~/[FilePath]/[FileName].las")

las$CRS #to check the CRS of the imported las. Returns 6340 -> EPSG code for NAD83(2011) / UTM zone 11N 

```

Notes regarding use of LAS Catalog format: 
- Efficient memory management, especially for large .las files
- Enables clipping of multifeature polygons (e.g. a shapefile containing multiple polygon features)

##Importing shapefile as simple feature (sf) class
```{r}
sf <- st_read(dsn = "C:\\Users\\shre9292\\OneDrive - University of Idaho\\Documents\\GitHub\\UAVShrubVolume\\DATA\\shrubs_arc.shp", layer = "shrubs_arc") 

sf #returns Dimensions as 'XY' and Projected CRS as NAD83(2011) / UTM zone 11N (same as point cloud on line 12), if sf has Dimensions XYZ, refer to NOT RUN section starting from line 68.

```

Notes regarding importing the shapfile for the case of REM 475 project: 
  - I imported the shapefile I created in ArcGIS where:
    - I imported the final output shapefile from Georgia's script (in EPSG 4152, geographic) into ArcGIS
    - Used the 'export' feature to export the EPSG 4152 geographic shapefile to EPSG 6340 projected shapefile

##Setting destination of the output clipped point cloud files (.las)
```{r}
opt_output_files(las) <- "~/[FilePath]/[FileName]_{ID}"

```

For e.g.: If '[FileName]_{ID}' is set as 'shrub_{Shrub_ID}', then the output file will be named as 'shrub_ARAR_1.las' where 'ARAR_1' is the unique ID ('{Shrub_ID}') of polygon feature of the shapefile, and hence is the ID of the individual shrub IRL.

##Clipping shapefile to point cloud 
```{r}
clipped_las <- clip_roi(las,sf)

```
There is no need to export the 'clipped_las' variable, the 'opt_output_files(las)' function automatically exports the files in the folder defined in line 55

#-------------------------------------------------------------------------------------------------------------------------------------------------------------
## NOT RUN: this section of the script can be used if the imported shapefile has XYZ dimension. This portion of the script removes the Z dimension and converts the sf to Spatial polygons class (SpP)
IF NEED TO RUN:

```{r}

#sf class has XYZ dimension, need to convert to 2 dimension (XY) 'SpatialPolygon' class:

sf_rmZ <- st_zm(sf) #removes Z dimension
SpP <- as(st_geometry(sf), "Spatial") #converts sf to sp class

```